Hunting experience shapes individual foraging specialization and predator-prey interactions in an online videogame

### 2023-03-09 manuscript/Chapter2-manuscript.Rmd
updated the discussion

The development of foraging tactics is crucial for young predators to reach adulthood and survive [@Phillips.etal2017;@Heithaus.etal2018]. Theory predicts that foraging specialization may emerge via learning, limitations in memorizing multiple complex hunting skills, and expertise [@Tinker.etal2009;@Dukas2019]. @Dukas2019 defines expertise as the characteristics, skills, and knowledge allowing individuals with extensive experience to outperform novices on complex tasks. This body of work suggests that the development of expertise through extensive practice is an optimizing process that promotes foraging specialization. Empirical studies on human and non-human hunters show that experience optimizes the efficiency (e.g. search and handling times, return rates) of their foraging tactics potentially via associative images or reliance on prey and environmental cues [@Edwards.Jackson1994;@Morse2000;@macdonaldCrossculturalComparisonLearning2007;@Reid.etal2010;@Wilson-Rankin2015]. Such optimization may thus reinforce the use of the same tactic (i.e. specialization) if its success is constant each time a prey is encountered. It may also be costly to attempt different hunting tactics by trial and error when prey are scarce or highly unpredictable [@Dukas1998;@Estes.etal2003;@Mery.Burns2010]. An alternative mechanism is that the costs/risks faced by hunters when switching foraging tactics may be offset by gaining experience and information on their prey, leading to increased individual foraging flexibility [@Stephens1993;@Ishii.Shimada2010;@Mery.Burns2010;@Kelley.Magurran2011;@Snell-Rood2013]. To develop proper responses to fluctuating resources, individuals would need to sample their environment broadly during an extensive period of time, followed by higher performance at later stages of development [reviewed in @Snell-Rood2013]. However, predators can differ in the challenges that they face during their lifetime, which could lead to both mechanisms operating at the same time within a predator population. For instance, some individuals may constantly encounter faster prey, which are proven to be more difficult to hunt [@Walker.etal2005;@Kelley.Magurran2011;@FraserFranco.etal2022].

@@

```{r fig2, out.width = "100%", fig.cap = "**Figure 2.** Distribution of the intra individual behavioural variation (i.e. specialization) of predators when they were (A) novices and then (B) advanced hunters. The red band displays the 95% CI of the intercept for the intra individual standard deviation in predator speed."}

@@

Hunting experience is predicted to imbue an increase in predator foraging specialization when resources are predictable [@Estes.etal2003;@Tinker.etal2008;@Tinker.etal2009;@Dukas2019]. Other lines of research view learning from experience as a form of behavioural plasticity [@Ishii.Shimada2010;@Mery.Burns2010;@Snell-Rood2013;@Wright.etal2022], with theoretical models predicting an increase in flexibility from experience when resources are variable [@Stephens1993]. Our results challenge the hypothesis that experience generates population-level behavioural responses. Instead, experience appears to be a property of individuals, such that each predator developed, through extensive practice, their own behavioural trajectory; some specialized at fast-paced hunting, while others adopted a slower and more flexible tactic. The predators' tactics were matched to prey encounters, such that specialized cursorial hunters experienced more predictable encounters, while slower and flexible hunters experienced more unpredictable encounters, with no difference in success between the two predator-types. This predator-prey behavioural matching grew stronger with predator experience, while the relationship between the prey's behaviour and hunting success diminished. Together, our observations suggest that predators learn to adjust their hunting tactic with experience, and the degree at which they specialize in it, based on the behaviour of the prey that they encountered throughout their lifetime. Moreover, the behavioural escalation through experience suggests that learning may be a catalyst for predator-prey arms races.

## How hunting experience shapes foraging behaviour

*- Individuals differ in the development of their expertise*
One of the main results of our study is that predators develop their own... (parler d'expertise et comportement)

*- Specialization does not change with experience at the population level*

*- Individuals either specialize or increase their flexiblity with experience*

This has important implications for future theoretical models predicting a directional increase in specialization or flexibility. should incorporate individual variation if both mechanisms operate at the same time.

**keep Warburton, 2003 in mind as a ref**
**Check Endler 1991 too from Kelley and Magurran**
**Try to find again the ref saying that high speed prey are more difficult**

## Consequences of hunting experience for predator-prey interactions

*- Experience strengthens predator-prey interactions*

@@

Our results also have implications also for predator-prey models of trait-matching, because such models are usually studied at the evolutionary scale.

*- Experience dampens the relationship between behaviour and success*

**Check if I should add the role of phenotypic matching in the intro, and the role of learning [added a sentence]**
**may add a sentence to better explain why prey difficulty is important. See Prokopenko and also Novak**
**garder en tête l'étude de Lima AmNat sur purposeful unpredictability**
**Iannou also found strong among individual differences in slopes of trial number**

>>>>

The development of foraging tactics is crucial for young predators to reach adulthood and survive [@Phillips.etal2017;@Heithaus.etal2018]. Theory predicts that foraging specialization may emerge via learning, limitations in memorizing multiple complex hunting skills, and expertise [@Tinker.etal2009;@Dukas2019]. @Dukas2019 defines expertise as the characteristics, skills, and knowledge allowing individuals with extensive experience to outperform novices on complex tasks. This body of work suggests that the development of expertise through extensive practice is an optimizing process that should promote foraging specialization. Empirical studies on human and non-human hunters show that experience optimizes the efficiency (e.g. search and handling times, return rates) of their foraging tactics potentially via associative images or reliance on prey and environmental cues [@Edwards.Jackson1994;@Morse2000;@macdonaldCrossculturalComparisonLearning2007;@Reid.etal2010;@Wilson-Rankin2015]. Such optimization may thus reinforce the use of the same tactic (i.e. specialization) if its success is constant each time a prey is encountered. It may also be costly to attempt different hunting tactics by trial and error when prey are scarce or highly unpredictable [@Dukas1998;@Estes.etal2003;@Mery.Burns2010]. An alternative mechanism is that the costs/risks faced by hunters when switching foraging tactics may be offset by gaining experience and information on their prey, leading to increased individual foraging flexibility [@Stephens1993;@Ishii.Shimada2010;@Mery.Burns2010;@Kelley.Magurran2011;@Snell-Rood2013]. To develop proper responses to fluctuating resources, individuals would need to sample their environment broadly during an extensive period of time, followed by higher performance at later stages of development [reviewed in @Snell-Rood2013]. However, predators can differ in the challenges that they face during their lifetime, which could lead to both mechanisms operating at the same time within a predator population. For instance, some individuals may constantly encounter faster prey, which are proven to be more difficult to hunt [@Walker.etal2005;@Kelley.Magurran2011;@FraserFranco.etal2022].

@@

```{r fig2, out.width = "100%", fig.cap = "**Figure 2.** Individual behavioural distribution of the predators' speed drawn from the MDHGLM estimates for individuals that had (A) the greatest increase in specialization and (B) the greatest increase in flexibility. The background (gray) distributions are for when individuals were novice hunters, and the foreground (blue) distributions with solid contour lines are for when they were advanced hunters."}

@@

Hunting experience is predicted to imbue an increase in predator foraging specialization when resources are predictable [@Estes.etal2003;@Tinker.etal2008;@Tinker.etal2009]. Other lines of research view learning from experience as a form of behavioural plasticity [@Ishii.Shimada2010;@Mery.Burns2010;@Snell-Rood2013;@Wright.etal2022], with theoretical models predicting an increase in flexibility from experience when resources are variable [@Stephens1993]. Our results challenge the hypothesis that experience should generate a directional population-level increase in behavioural specialization or plasticity. Instead, the consequences from experience appear to be a property of individuals, such that each predator developed, through extensive practice, their own expertise and behavioural trajectory; some specialized at fast-paced hunting, while others adopted a slower and more flexible tactic. The predators' tactics were matched to prey encounters, such that specialized cursorial hunters experienced more predictable encounters, while slower and flexible hunters experienced more unpredictable encounters, with no difference in success between the two predator-types. This predator-prey behavioural matching grew stronger with predator experience, while the relationship between the prey's behaviour and hunting success diminished. Together, our observations suggest that predators learn to adjust their hunting tactic with experience, and the degree at which they specialize in it, based on the behaviour of the prey that they encountered throughout their lifetime. Moreover, the behavioural escalation through experience suggests that learning may be a catalyst for predator-prey arms races.

## The development of expertise with hunting experience

Empirical studies in humans and animals show that proficiency in a given task often increases nonlinearly with experience, and stabilizes when an expert level is reached [reviewed in @Dukas2019]. However, these patterns are derived from models that assume the same increase in expertise for each individual, and their data do not include repeated measures of individual success through time. In contrast with these studies, our analyses revealed that there was no population increase in expertise; instead, individual predators in *DBD* displayed distinct patterns of expertise acquisition. For instance, many predators increased their success (Figure S1A), but others displayed a decrease in success (Figure S1B) or no change in success at all with experience. We predicted that this would occur because predators should differ in the type of prey that they encounter, as some may be easier to capture than others. This is often the case in nature as prey can use variety of defenses such as physical armaments, toxins, camouflage, or antipredator behaviours to escape predation [@BrodieIII.BrodieJr.1999;@Bowen.etal2002;@brodieCoEvolutionPredatorsPrey2010;@Nomura.etal2011;@Carey.Wahl2011]. However, our results show that the speed of the prey did not shape individual differences in skill aquisition by predators, even if they differed in the prey that they encountered. This was unexpected, because prey speed is an important predictor of predator success in this system [@FraserFranco.etal2022] as well as in others [@Walker.etal2005;@Kelley.Magurran2011]. Despite that other antipredator tactics may mediate these individual responses by predators, it is likely that individual differences in motivation and perseverance, attention span, or memory retention were also major causes [@Warburton2003;@morand-ferronStudyingEvolutionaryEcology2016;@Dukas2019]. For instance, a persevering predator may increase its success much more steadily than an unmotivated predator that quickly abandons a chase. Another factor that may explain differences in the development of expertise is the time delay between hunting events, such that a predator that played 300 matches in the span of six months may not retain information (e.g. prey escape patterns or muscle memory) as well as one that played 300 matches in the span of 6 days. While this has not been tested yet in a predator-prey interaction, it is hypothesized that a greater delay between exposure events should attenuate or even negate the relationship between cumulative experience and success as a result of forgetting [@Endler1991;@Wright.etal2022].

## Changes in individual hunting specialization with experience and consequences for predator-prey interactions

We detected a similar pattern for foraging specialization, such that individual differences increased with experience, while specialization at the population level remained stable. *parler du patron non linéaire et donner quelques potentielles explications liées à l'exploration + grande chez les juvéniles*. *Je peux citer les sources de Phillips et al : Thiers 2014 et Grissac 2016*.

Predators that experienced predictable encounters with their prey specialized in hunting at high speeds, while those that experienced more unpredictable encounters adopted a slower and flexible strategy. This is similar to other studies finding that fast-paced hunting is a highly specialized tactic suited for prey that use rapid escapes [@Endler1991;@Bro-Jorgensen2013;@Wilson.etal2018;@Szopa-Comley.Ioannou2022]. On the contrary, instead of focusing solely on one tactic, predators that experienced unpredictable encounters probably adjusted their behaviour to minimize the consequences of uncertainty. Interestingly, both specialist and flexible hunters fine-tuned their tactic to their prey as they gained experience, suggesting that they learned how to hunt their prey. In that regard, our results provide direct evidence that the mechanism underlying the resource-predictability hypothesis may indeed involve predators learning to adjust their degree of specialization based on their prey [@Weimerskirch.etal2005;@Weimerskirch2007;@Woo.etal2008;@Phillips.etal2017]. This is further supported by the fact that specialist and flexible hunters both achieved similar success throughout the study period. Therefore, our results have important implications for future predator-prey models, which should incorporate individual variation in learning and prey encounters to better predict under which ecological contexts specialization should be favoured over flexibility.

*je peux commencer à plug le arms race dans ce paragraphe.*
*il faut aussi que je parle du fait que la variation dans le succès de capture a diminué avec l'XP*
*- Experience strengthens predator-prey interactions*
*- Experience dampens the relationship between behaviour and success*

@@

**garder en tête l'étude de Lima AmNat sur purposeful unpredictability**
**Iannou also found strong among individual differences in slopes of trial number**

*Our results also have implications also for predator-prey models of trait-matching, because such models are usually studied at the evolutionary scale.*
